---
title: 'Developing Data Products Course Project: Shiny Application and Reproducible
  Pitch'
author: "suchit malhotra"
date: "May 20, 2018"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Overview

This application gives you a glimpse of Swiss Data set and corelation between various parameters.

Thi shiny application is developed for this assignment is available at: <https://suchitmalhotra.shinyapps.io/shinyProject/>

The source code of ui.R and server.R is available on Github repo at: <https://github.com/malhotrasuchit/ProjectAssignment-DataScience/tree/master/Developing%20Data%20Products/Week%204/shinyProject>

## Swiss Dataset

Swiss Fertility and Socioeconomic Indicators (1888) Data

Description

Standardized fertility measure and socio-economic indicators for each of 47 French-speaking provinces of Switzerland at about 1888.

Details

Switzerland, in 1888, was entering a period known as the demographic transition; i.e., its fertility was beginning to fall from the high level typical of underdeveloped countries.

The data collected are for 47 French-speaking "provinces" at about 1888.

Here, all variables are scaled to [0, 100], where in the original, all but "Catholic" were scaled to [0, 1].

## Shiny App: Exploring Swiss Dataset

Type of plot

- Plot: Simple Plot depicting Fertility ~ Infant.Mortality

- ggpairs: depicts a ggplot2 generalized pairs plot

- pairs: depicts a matrix of scatterplot of swiss dataset

## Swiss Dataset Plots

```{r message=FALSE, warning=FALSE}
library(GGally)
plot(swiss$Fertility, swiss$Infant.Mortality, data = swiss, col= swiss$Catholic)
pairs(swiss, panel = panel.smooth, main = "swiss data",col = 3 + (swiss$Catholic > 50))
ggpairs(swiss)
```

