---
title: "Interactive Data Mapping using Leafleat - Data Products"
date: "May 16, 2018"
output: html_document
---

##Peer-graded Assignment: R Markdown and Leaflet

Create a web page using R Markdown that features a map created with Leaflet.

Host your webpage on either GitHub Pages, RPubs, or NeoCities.

Your webpage must contain the date that you created the document, and it must contain a map created with Leaflet. We would love to see you show off your creativity

##Solution

Load Leaflet library
```{r}
library(leaflet)
```

By accessing the Open Data NYC api we can look at art museum locations in New York City.

```{r}
library(jsonlite)
data <- fromJSON("https://data.cityofnewyork.us/resource/43hw-uvdj.json", flatten = TRUE)
```

``` {r}
artLatLong <- data.frame(
  lat = unlist(lapply(data$the_geom.coordinates, `[[`, 2)),
  lng = unlist(lapply(data$the_geom.coordinates, `[[`, 1))
)

artIcons <- makeIcon(
  iconUrl = "photo.png",
  iconWidth = 35*215/230, iconHeight = 35,
  iconAnchorX = 35*215/230/2, iconAnchorY = 35
)

artSites <- paste("<a href='", data$url, "'>",data$name,"</a>" ,sep = "")
```

## May 16, 2018
###Map of Art Museums in NYC
``` {r}
artLatLong %>% leaflet() %>% addTiles() %>%
addMarkers(icon=artIcons, popup = artSites)
```

